Methods to prioritise pop-up active transport infrastructure

### 2020-05-10 article.Rmd
Justify appropriateness of 5 mile buffer
Note limitations on use of lane count data, as a basis for future work
Note that method is reproducible beyond the UK, as OSM is an international dataset

<!-- The region of analysis may seem like a basic consideration: most cities have well-defined administrative zones. -->
<!-- In Leeds and many other cities, it makes sense to focus on the region directly surrounding the city centre, in a kind of 'geographical triage' to omit from the analysis pop-up options in the outskirts and focus valuable attention on the routes that are most likely to serve the highest number of people. -->
We set a modifiable parameter `city_centre_buffer_radius` with an initial value of 8 km (5 miles) to geographically subset potential routes.
Figure \@ref(fig:gsub) shows the result of subsetting based on physical distance from the centre vs plotting all possible transport network segments within the city boundaries.
To ensure roads that could serve key destinations were included, the parameter `key_destination_buffer_radius` (initially set to 5 km) was used.
<!-- could say more... -->

@@

To identify road sections on which there is a spare lane we developed a simple algorithm that takes the OSM variable [`lanes`](https://wiki.openstreetmap.org/wiki/Key:lanes)  if it is present and, if not, derives the number from the highway type and presence/absence of bus lanes.
All segments defined as having a spare lane using this method is shown in Figure \@ref(fig:levels) (left).
The result of filtering by distance and cycling potential before and after grouping using graph membership of touching roads is shown in \@ref(fig:levels) (middle and right, respectively).
Grouping linked roads before filtering results in a more cohesive network.

@@

The approach is not without limitations.
Its reliance on data rather than community engagement represents a rather top-down approach to transport planning.
To overcome this issue future work could seek to incorporate the results such as those presented above into a participatory map of the type shown in Figure 1.
Further work could also extend the method in various ways, for example by refining estimates of cycling potential based on new parameters such as proximity to key destinations.

A major advantage of the approach is that it is scalable.
It would be feasible to run the method for every city in the UK for which there is data, given sufficient computer and developer resource.
Given the recent interest in and funding for pop-up cycleways, rolling-out the method quickly, while being agile to adapt the method and parameters for different cities, could help ensure that funding for pop-up infrastructure is spent in an evidence-based way.

>>>>

<!-- The region of analysis may seem like a basic consideration: most cities have well-defined administrative zones. -->
<!-- In Leeds and many other cities, it makes sense to focus on the region directly surrounding the city centre, in a kind of 'geographical triage' to omit from the analysis pop-up options in the outskirts and focus valuable attention on the routes that are most likely to serve the highest number of people. -->
We set a modifiable parameter `city_centre_buffer_radius` with an initial value of 8 km (5 miles) to geographically subset potential routes. 5 miles represents a distance that most people have the physical ability to cycle.
Figure \@ref(fig:gsub) shows the result of subsetting based on physical distance from the centre vs plotting all possible transport network segments within the city boundaries.
To ensure roads that could serve key destinations were included, the parameter `key_destination_buffer_radius` (initially set to 5 km) was used.
<!-- could say more... -->

@@

To identify road sections on which there is a spare lane we developed a simple algorithm that takes the OSM variable [`lanes`](https://wiki.openstreetmap.org/wiki/Key:lanes)  if it is present and, if not, derives the number from the highway type and presence/absence of bus lanes.
All segments defined as having a spare lane using this method is shown in Figure \@ref(fig:levels) (left).
In future, this methodology could be enhanced to take into account the effect of lane widths, which increase the effective available space, and the presence of vehicle parking bays, which reduce it.
The result of filtering by distance and cycling potential before and after grouping using graph membership of touching roads is shown in \@ref(fig:levels) (middle and right, respectively).
Grouping linked roads before filtering results in a more cohesive network.

@@

The approach is not without limitations.
Its reliance on data rather than community engagement represents a rather top-down approach to transport planning.
To overcome this issue, future work could seek to incorporate the results such as those presented above into a participatory map of the type shown in Figure 1.
Further work could also extend the method in various ways, for example by refining estimates of cycling potential based on new parameters such as proximity to key destinations.

A major advantage of the approach is that it is scalable.
It would be feasible to run the method for every city in the UK (and indeed beyond) for which there is data, given sufficient computer and developer resource.
Given the recent interest in and funding for pop-up cycleways, rolling-out the method quickly, while being agile to adapt the method and parameters for different cities, could help ensure that funding for pop-up infrastructure is spent in an evidence-based way.
